---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# spotifyradar

<!-- badges: start -->

[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)

<!-- badges: end -->

`spotifyradar` is a shiny app that continuously summarizes my listening habits. I use [IFTTT](https://ifttt.com/) to upload basic track information to a Google sheet, which I monitor daily. I use [spotifyr](https://www.rcharlie.com/spotifyr/) to obtain more metadata for the tracks and artists I listen to. Then, I summarize everything in an explorable Shiny app.

In addition to my interest in my own listening habits outside of Spotify's end of year summary, I'm using this app as a tool to learn how to stream data and maintain code and data that is updated regularly. In addition, I'm practicing my SQL skills with data management and extraction.

## Setup

### Prior to installation

Before installing `spotifyradar`, you need to create an IFTTT account at [ifttt.com](ifttt.com). Then, add this applet to your account: [spotifyradar](https://ifttt.com/applets/UzkBGfpb-if-a-recently-played-track-then-add-row-to-my-google-drive-spreadsheet). The applet monitors your Spotify plays and updates a Google Sheet regularly with some basic info about the track. Your `spotifyradar` summary gets more interesting as you collect more data about your listening.

In addition, you need to set up a [Spotify Dev account](https://developer.spotify.com/my-applications/#!/applications). `spotifyradar` uses the [spotifyr](https://www.rcharlie.com/spotifyr/) wrapper around the Spotify API to obtain more info about tracks and artists, which requires Dev access.

After setting up your Dev account, you need to add your `SPOTIFY_CLIENT_ID` and `SPOTIFY_CLIENT_SECRET` to a `.Renviron`. The easiest way to do this is with the [usethis](https://usethis.r-lib.org/) function `usethis::edit_r_environ()`, which opens the .Renviron file and allows you to edit the file directly. Alternatively, you can create a blank `.Renviron` file in the project directory and use your favorite text editor to edit the file. The environmental variables should be specified in the `.Renviron` file like this:

    SPOTIFY_CLIENT_ID='xxxxxxxx'
    SPOTIFY_CLIENT_SECRET='xxxxxxx'

where `xxxxxxxx` contains your ID or SECRET. Restart R to make them available.

### Installation

To install `spotifyradar`:

```{r, eval=FALSE}
# if you don't have the remotes package already
install.packages(remotes)

remotes::install_github("connor-french/spotifyradar")
```

### Cron scheduling

`spotifyradar` uses `cronR` to schedule regular downloads of your Spotify data. The `cron_setup.R` script in the `R` folder is included as a convenience to setup the data streaming. Run this script and your data streaming is good to go! For this to work, you need to set your `SPOTIFY_CLIENT_ID` and `SPOTIFY_CLIENT_SECRET` environment variables. See the **Prior to installation** section for instructions.
